Global Deaths Due to Air Pollution

### 2022-05-01 outline.Rmd
Pollution types are done

death_17_low<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2017, country %in% c('United States', 'Brazil','Nigeria', 'Germany', 'Pakistan', 'Australia')) %>%
  select(total_deaths) %>% 
  summarize(Low_Deaths_17 = mean(total_deaths))

@@

Let's graph the previous tables!
```{r, echo=FALSE}
#High Population Deaths 2007 and 20012
p_high <- all_high_second %>%
  group_by(country) %>% 
  ggplot(aes(x= High_Deaths_07, y= High_Deaths_12, color= High_Deaths_17)) +
  geom_point() +
  facet_wrap(~country)

interact_high<- p_high + labs(title="High Populated Countries", x="2007 Deaths", y="2012 Deaths")

ggplotly(interact_high)

#Low Population Deaths 2007 and 2012
p_low <- all_low_second %>%
  group_by(country) %>% 
  ggplot(aes(x= Low_Deaths_07, y= Low_Deaths_12, color= Low_Deaths_17)) +
  geom_point() +
  facet_wrap(~country)

interact_low<- p_low + labs(title="Low Populated Countries", x="2007 Deaths", y="2012 Deaths")

ggplotly(interact_low)

@@

## Which year had the worst indoor? Outdoor particulate? Outdoor ozone? 
# Which is worse - outdoor or indoor pollution?

First, we split the data into high and low population based on country

@@

>>>>

death_17_low<- deaths_df %>% 
  group_by(country) %>% 
  filter(year == 2017, country %in% c('Canada', 'Chile', 'Malawi', 'Serbia', 'Sri Lanka', 'Tonga')) %>%
  select(total_deaths) %>% 
  summarize(Low_Deaths_17 = mean(total_deaths))

@@

Let's graph the previous tables!

The first decade.
```{r, echo=FALSE}
#High Population Deaths 1996-2006 
p_high_first <- all_high_first %>%
  group_by(country) %>% 
  ggplot(aes(x= High_Deaths_96, y= High_Deaths_01, color= High_Deaths_06)) +
  geom_point() +
  facet_wrap(~country)

interact_high_first<- p_high_first + labs(title="High Populated Countries", x="1996 Deaths", y="2001 Deaths")

ggplotly(interact_high_first)

#Low Population Deaths 1996-2006
p_low_first <- all_low_first %>%
  group_by(country) %>% 
  ggplot(aes(x= Low_Deaths_96, y= Low_Deaths_01, color= Low_Deaths_06)) +
  geom_point() +
  facet_wrap(~country)

interact_low_first<- p_low_first + labs(title="Low Populated Countries", x="1996 Deaths", y="2001 Deaths")

ggplotly(interact_low_first)
```

This shows the second decade.
```{r, echo=FALSE}
#High Population Deaths 2007-2017
p_high_second <- all_high_second %>%
  group_by(country) %>% 
  ggplot(aes(x= High_Deaths_07, y= High_Deaths_12, color= High_Deaths_17)) +
  geom_point() +
  facet_wrap(~country)

interact_high_second<- p_high_second + labs(title="High Populated Countries", x="2007 Deaths", y="2012 Deaths")

ggplotly(interact_high_second)

#Low Population Deaths 2007-2017
p_low_second <- all_low_second %>%
  group_by(country) %>% 
  ggplot(aes(x= Low_Deaths_07, y= Low_Deaths_12, color= Low_Deaths_17)) +
  geom_point() +
  facet_wrap(~country)

interact_low_second<- p_low_second + labs(title="Low Populated Countries", x="2007 Deaths", y="2012 Deaths")

ggplotly(interact_low_second)

@@

> Which year had the worst indoor? Outdoor particulate? Outdoor ozone?

Indoor Deaths
```{r, echo=FALSE}
#Indoor 1996-2017
indoor_high<- deaths_df %>% 
    filter(year> 1995 & country %in% c('United States', 'Brazil','Nigeria', 
                                       'Germany', 'Pakistan', 'Australia')) %>% 
  group_by(country, year, indoor_deaths) %>% 
  ggplot()+
  geom_point(aes(year, indoor_deaths, color = country))+
  ggtitle("High Populated Indoor Death 1996-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(indoor_high)

indoor_low<- deaths_df %>% 
    filter(year> 1995 & country %in% c('Canada', 'Chile', 'Malawi', 
                                       'Serbia', 'Sri Lanka', 'Tonga')) %>% 
  group_by(country, year, indoor_deaths) %>% 
  ggplot()+
  geom_point(aes(year, indoor_deaths, color = country))+
  ggtitle("Low Populated Indoor Death 1996-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(indoor_low)

#Pakistan and Malawi were both mainly affected in 1996
indoor_top<- deaths_df %>% 
    filter(year> 1995 & country %in% c('Pakistan', 'Malawi')) %>% 
  group_by(country, year, indoor_deaths) %>% 
  ggplot()+
  geom_point(aes(year, indoor_deaths, color = country))+
  ggtitle("Greatest Indoor Death 1996-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(indoor_top)
```

Outdoor Deaths
```{r, echo= FALSE}
#Outdoor Particulate 1996-2017
outdoor_high<- deaths_df %>% 
    filter(year> 1995 & country %in% c('United States', 'Brazil','Nigeria', 
                                       'Germany', 'Pakistan', 'Australia')) %>% 
  group_by(country, year, outdoor_deaths) %>% 
  ggplot()+
  geom_point(aes(year, outdoor_deaths, color = country))+
  ggtitle("High Populated Outdoor Death 1996-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(outdoor_high)

outdoor_low<- deaths_df %>% 
    filter(year> 1995 & country %in% c('Canada', 'Chile', 'Malawi', 
                                       'Serbia', 'Sri Lanka', 'Tonga')) %>% 
  group_by(country, year, outdoor_deaths) %>% 
  ggplot()+
  geom_point(aes(year, outdoor_deaths, color = country))+
  ggtitle("Low Populated Outdoor Death 1996-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(outdoor_low)

#Serbia and Pakistan were the top countries affected. 
#2011 Pakistan had the greatest amount of Outdoor Deaths
outdoor_top<- deaths_df %>% 
    filter(year> 1995 & country %in% c('Pakistan', 'Serbia')) %>% 
  group_by(country, year, outdoor_deaths) %>% 
  ggplot()+
  geom_point(aes(year, outdoor_deaths, color = country))+
  ggtitle("Greatest Outdoor Death 1996-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(outdoor_top)
```

***
Ozone Deaths 
```{r, echo=FALSE}
#Outdoor ozone 1996-2017
ozone_high<- deaths_df %>% 
    filter(year> 1995 & country %in% c('United States', 'Brazil','Nigeria', 
                                       'Germany', 'Pakistan', 'Australia')) %>% 
  group_by(country, year, ozone_deaths) %>% 
  ggplot()+
  geom_point(aes(year, ozone_deaths, color = country))+
  ggtitle("High Populated Ozone Death 1996-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(ozone_high)

ozone_low<- deaths_df %>% 
    filter(year> 1995 & country %in% c('Canada', 'Chile', 'Malawi', 
                                       'Serbia', 'Sri Lanka', 'Tonga')) %>% 
  group_by(country, year, ozone_deaths) %>% 
  ggplot()+
  geom_point(aes(year, ozone_deaths, color = country))+
  ggtitle("Low Populated Ozone Death 1996-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(ozone_low)

#Malawi and Pakistan were the top countries affected 
#1997 had the greatest amount of ozone death in Pakistan
ozone_top<- deaths_df %>% 
    filter(year> 1995 & country %in% c('Pakistan', 'Malawi')) %>% 
  group_by(country, year, ozone_deaths) %>% 
  ggplot()+
  geom_point(aes(year, ozone_deaths, color = country))+
  ggtitle("Greatest Ozone Death 1996-2017")+
  xlab("Year")+
  ylab("Deaths per 100,000")
ggplotly(ozone_top)
```

# Which is worse? 
> outdoor or indoor pollution?
```{r}

# Reminder: Include this in the beginning 
First, we split the data into high and low population based on country

@@

# Include Earlier to See who's been affected
